---
title: "Cyclistic_tripdata_RM"
output: html_document
author: Samuel Johnson
date: 01/03/2022
---

## Extract Data from Files

Downloaded orignial files of previous 12 monthes of trip data from:

https://divvy-tripdata.s3.amazonaws.com/index.html


During the main part of the data cleaning process, I chose to utilize pipes in order to keep my code more condensed and easier to read. The drop_na() function was used to get rid of null values. The select() function was used to pick out only the columns of interest (rideable_type, start_station_name, start_station_id, end_station_name, end_station_id, and member_casual). 

I used the following code to generate 12 cleaned data frames simply by altering the dataset id (yyyymm) in each line, this was done for datasets December-2020 (202012) through November-2021 (202111).

```{r}
library("tidyverse")
library("janitor")



X202012_DF <- read_csv("202012-divvy-tripdata.csv")

cleaned_202012_DF <- X202012_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202012_DF,"CleanedBikeData202012.csv", row.names = FALSE)




# Dataset_id_202101

X202101_DF <- read_csv("202101-divvy-tripdata.csv")

cleaned_202101_DF <- X202101_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202101_DF,"CleanedBikeData202101.csv", row.names = FALSE)




# Dataset_id_202102

X202102_DF <- read_csv("202102-divvy-tripdata.csv")

cleaned_202102_DF <- X202102_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202102_DF,"CleanedBikeData202102.csv", row.names = FALSE)




# Dataset_id_202103

X202103_DF <- read_csv("202103-divvy-tripdata.csv")

cleaned_202103_DF <- X202103_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202103_DF,"CleanedBikeData202103.csv", row.names = FALSE)




# Dataset_id_202104

X202104_DF <- read_csv("202104-divvy-tripdata.csv")

cleaned_202104_DF <- X202104_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202104_DF,"CleanedBikeData202104.csv", row.names = FALSE)




# Dataset_id_202105

X202105_DF <- read_csv("202105-divvy-tripdata.csv")

cleaned_202105_DF <- X202105_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202105_DF,"CleanedBikeData202105.csv", row.names = FALSE)




# Dataset_id_202106

X202106_DF <- read_csv("202106-divvy-tripdata.csv")

cleaned_202106_DF <- X202106_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202106_DF,"CleanedBikeData202106.csv", row.names = FALSE)




# Dataset_id_202107

X202107_DF <- read_csv("202107-divvy-tripdata.csv")

cleaned_202107_DF <- X202107_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202107_DF,"CleanedBikeData202107.csv", row.names = FALSE)




# Dataset_id_202108

X202108_DF <- read_csv("202108-divvy-tripdata.csv")

cleaned_202108_DF <- X202108_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202108_DF,"CleanedBikeData202108.csv", row.names = FALSE)




# Dataset_id_202109

X202109_DF <- read_csv("202109-divvy-tripdata.csv")

cleaned_202109_DF <- X202109_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202109_DF,"CleanedBikeData202109.csv", row.names = FALSE)




# Dataset_id_202110

X202110_DF <- read_csv("202110-divvy-tripdata.csv")

cleaned_202110_DF <- X202110_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202110_DF,"CleanedBikeData202110.csv", row.names = FALSE)




# Dataset_id_202111

X202111_DF <- read_csv("202111-divvy-tripdata.csv")

cleaned_202111_DF <- X202111_DF %>% 
  drop_na() %>% 
  select("member_casual", "rideable_type", "start_station_name", "start_station_id", "end_station_name", "end_station_id") 
  
write.csv(cleaned_202111_DF,"CleanedBikeData202111.csv", row.names = FALSE)


```

## Get Initial Totals for Each Month

The summarize() and sum() functions were used in conjunction to get totals for annual members, casual users, total users, docked bikes, electric bikes, classic bikes, and total bikes for each individual, cleaned dataset.

I entered these results into a new spreadsheet in order to make pivot tables. I used the condition function =IF(E2=O2, TRUE()) to check that my data had been entered correctly and that the total number of users was equal to the total number of bikes for each individual dataset.

```{r}

# Dataset_id_202012

trim202012_DF <- cleaned_202012_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202012_DF)


docked_202012_DF <- cleaned_202012_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202012_DF)


electric_202012_DF <- cleaned_202012_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202012_DF)


classic_202012_DF <- cleaned_202012_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202012_DF)



# Dataset_id_202101

trim202101_DF <- cleaned_202101_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202101_DF)


docked_202101_DF <- cleaned_202101_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202101_DF)


electric_202101_DF <- cleaned_202101_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202101_DF)


classic_202101_DF <- cleaned_202101_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202101_DF)



# Dataset_id_202102

trim202102_DF <- cleaned_202102_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202102_DF)


docked_202102_DF <- cleaned_202102_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202102_DF)


electric_202102_DF <- cleaned_202102_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202102_DF)


classic_202102_DF <- cleaned_202102_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202102_DF)



# Dataset_id_202103

trim202103_DF <- cleaned_202103_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202103_DF)


docked_202103_DF <- cleaned_202103_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202103_DF)


electric_202103_DF <- cleaned_202103_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202103_DF)


classic_202103_DF <- cleaned_202103_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202103_DF)




# Dataset_id_202104

trim202104_DF <- cleaned_202104_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202104_DF)


docked_202104_DF <- cleaned_202104_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202104_DF)


electric_202104_DF <- cleaned_202104_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202104_DF)


classic_202104_DF <- cleaned_202104_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202104_DF)



# Dataset_id_202105

trim202105_DF <- cleaned_202105_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202105_DF)


docked_202105_DF <- cleaned_202105_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202105_DF)


electric_202105_DF <- cleaned_202105_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202105_DF)


classic_202105_DF <- cleaned_202105_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202105_DF)



# Dataset_id_202106

trim202106_DF <- cleaned_202106_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202106_DF)


docked_202106_DF <- cleaned_202106_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202106_DF)


electric_202106_DF <- cleaned_202106_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202106_DF)


classic_202106_DF <- cleaned_202106_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202106_DF)



# Dataset_id_202107

trim202107_DF <- cleaned_202107_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202107_DF)


docked_202107_DF <- cleaned_202107_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202107_DF)


electric_202107_DF <- cleaned_202107_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202107_DF)


classic_202107_DF <- cleaned_202107_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202107_DF)



# Dataset_id_202108

trim202108_DF <- cleaned_202108_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202108_DF)


docked_202108_DF <- cleaned_202108_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202108_DF)


electric_202108_DF <- cleaned_202108_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202108_DF)


classic_202108_DF <- cleaned_202108_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202108_DF)



# Dataset_id_202109

trim202109_DF <- cleaned_202109_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202109_DF)


docked_202109_DF <- cleaned_202109_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202109_DF)


electric_202109_DF <- cleaned_202109_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202109_DF)


classic_202109_DF <- cleaned_202109_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202109_DF)



# Dataset_id_202110

trim202110_DF <- cleaned_202110_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202110_DF)


docked_202110_DF <- cleaned_202110_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202110_DF)


electric_202110_DF <- cleaned_202110_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202110_DF)


classic_202110_DF <- cleaned_202110_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202110_DF)




# Dataset_id_202111

trim202111_DF <- cleaned_202111_DF %>%
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(trim202111_DF)


docked_202111_DF <- cleaned_202111_DF %>%
  filter(rideable_type == "docked_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(docked_202111_DF)


electric_202111_DF <- cleaned_202111_DF %>%
  filter(rideable_type == "electric_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(electric_202111_DF)


classic_202111_DF <- cleaned_202111_DF %>%
  filter(rideable_type == "classic_bike") %>% 
  summarize(num_of_member = sum(member_casual == "member"), num_of_casual = sum(member_casual == "casual"))

View(classic_202111_DF)


```

## Merge Datasets

At this point, the 12 cleaned datasets were stitched together using the bind_rows() and list() functions. This created 1 complete and cleaned data frame.

```{r}

CleanedBikeData202012 <- read_csv("CleanedBikeData202012.csv")

CleanedBikeData202101 <- read_csv("CleanedBikeData202101.csv")

CleanedBikeData202102 <- read_csv("CleanedBikeData202102.csv")

CleanedBikeData202103 <- read_csv("CleanedBikeData202103.csv")

CleanedBikeData202104 <- read_csv("CleanedBikeData202104.csv")

CleanedBikeData202105 <- read_csv("CleanedBikeData202105.csv")

CleanedBikeData202106 <- read_csv("CleanedBikeData202106.csv")

CleanedBikeData202107 <- read_csv("CleanedBikeData202107.csv")

CleanedBikeData202108 <- read_csv("CleanedBikeData202108.csv")

CleanedBikeData202109 <- read_csv("CleanedBikeData202109.csv")

CleanedBikeData202110 <- read_csv("CleanedBikeData202110.csv")

CleanedBikeData202111 <- read_csv("CleanedBikeData202111.csv")

Complete_BikeData_DF <- bind_rows(list(CleanedBikeData202012, CleanedBikeData202101, CleanedBikeData202102, CleanedBikeData202103, CleanedBikeData202104, CleanedBikeData202105, CleanedBikeData202106, CleanedBikeData202107, CleanedBikeData202108, CleanedBikeData202109, CleanedBikeData202110, CleanedBikeData202111))

write.csv(Complete_BikeData_DF,"CompleteCleanedBikeData.csv", row.names = FALSE)

```

## Find Most Used Routes

To find the most used routes, I used the count() function followed by the arrange() function in descending order. Used the mutate(), round(), sum(), and paste0() functions to find the percentage that route was used. The slice() function was used to limit it to the top 10 most used routes.

The filter() function was used to find the most used routes by members vs casual-users.

```{r}


# Get list of most used stations (member & casual)

library("tidyverse")

entire_df <- read_csv("CompleteCleanedBikeData.csv")

top_routes_df <- entire_df %>% 
  count(start_station_name, end_station_name) %>% 
  arrange(-n) %>% 
  mutate(
    pct = round(n / sum(n) * 100, 3),
    pct_fmt = paste0(pct, "%")
  ) %>% 
  slice(1:10)

write.csv(top_routes_df,"TopRoutesGeneral_.csv", row.names = FALSE)





# Get list of most used stations (member only)

library("tidyverse")

entire_member_df <- read_csv("CompleteCleanedBikeData.csv") %>% 
  filter(member_casual == "member")

top_routes_member_df <- entire_member_df %>%
  count(start_station_name, end_station_name) %>% 
  arrange(-n) %>% 
  mutate(
    pct = round(n / sum(n) * 100, 3),
    pct_fmt = paste0(pct, "%")
  ) %>% 
  slice(1:10)

write.csv(top_routes_member_df,"TopRoutesMember_.csv", row.names = FALSE)





# Get list of most used stations (casual only)

library("tidyverse")

entire_casual_df <- read_csv("CompleteCleanedBikeData.csv") %>% 
  filter(member_casual == "casual")

top_routes_casual_df <- entire_casual_df %>%
  count(start_station_name, end_station_name) %>% 
  arrange(-n) %>% 
  mutate(
    pct = round(n / sum(n) * 100, 3),
    pct_fmt = paste0(pct, "%")
  ) %>% 
  slice(1:10)

write.csv(top_routes_casual_df,"TopRoutesCasual_.csv", row.names = FALSE)


```

## Create Bar Charts


Next, I created bar charts for 5 of the most popular start stations for casual-users as well as 1 bar chart that included data from all start stations. The x-axis is "Bike-Type" (rideable_type), the y-axis is the number of starts (count), and the charts are faceted by "Rider Type" (member_casual). 

Start Station Names:

Streeter Dr & Grand Ave
Michigan Ave & Oak St
Millennium Park
Lake Shore Dr & Monroe St
Buckingham Fountain

The following code was used to generate bar charts for the 5 start stations listed above.

```{r}

# "Streeter Dr & Grand Ave" Bar Chart

library("tidyverse")

entire_df <- read_csv("CompleteCleanedBikeData.csv") %>% 
  filter(start_station_name == "Streeter Dr & Grand Ave")

ggplot(data = entire_df)+
  geom_bar(mapping = aes(x=rideable_type, fill=member_casual))+
  labs(
    title = "Starts per Bike-Type: Streeter Dr & Grand Ave", 
    subtitle = "Dec 2020 - Nov 2021",
    fill = "Rider type",
    x = "Bike-Type"
  )+
  scale_fill_manual(values = c("#339999","#FF0000"))+
  facet_wrap(~member_casual)+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 30),
    panel.grid.major.x = element_blank()
  )




# "Millennium Park" Bar Chart

library("tidyverse")

entire_df <- read_csv("CompleteCleanedBikeData.csv") %>% 
  filter(start_station_name == "Millennium Park")

ggplot(data = entire_df)+
  geom_bar(mapping = aes(x=rideable_type, fill=member_casual))+
  labs(
    title = "Starts per Bike-Type: Millennium Park", 
    subtitle = "Dec 2020 - Nov 2021",
    fill = "Rider type",
    x = "Bike-Type"
  )+
  scale_fill_manual(values = c("#339999","#FF0000"))+
  facet_wrap(~member_casual)+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 30),
    panel.grid.major.x = element_blank()
  )




# "Michigan Ave & Oak St" Bar Chart

library("tidyverse")

entire_df <- read_csv("CompleteCleanedBikeData.csv") %>% 
  filter(start_station_name == "Michigan Ave & Oak St")

ggplot(data = entire_df)+
  geom_bar(mapping = aes(x=rideable_type, fill=member_casual))+
  labs(
    title = "Starts per Bike-Type: Michigan Ave & Oak St", 
    subtitle = "Dec 2020 - Nov 2021",
    fill = "Rider type",
    x = "Bike-Type"
  )+
  scale_fill_manual(values = c("#339999","#FF0000"))+
  facet_wrap(~member_casual)+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 30),
    panel.grid.major.x = element_blank()
  )




# "Lake Shore Dr & Monroe St" Bar Chart

library("tidyverse")

entire_df <- read_csv("CompleteCleanedBikeData.csv") %>% 
  filter(start_station_name == "Lake Shore Dr & Monroe St")

ggplot(data = entire_df)+
  geom_bar(mapping = aes(x=rideable_type, fill=member_casual))+
  labs(
    title = "Starts per Bike-Type: Lake Shore Dr & Monroe St", 
    subtitle = "Dec 2020 - Nov 2021",
    fill = "Rider type",
    x = "Bike-Type"
  )+
  scale_fill_manual(values = c("#339999","#FF0000"))+
  facet_wrap(~member_casual)+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 30),
    panel.grid.major.x = element_blank()
  )




# "Buckingham Fountain" Bar Chart

library("tidyverse")

entire_df <- read_csv("CompleteCleanedBikeData.csv") %>% 
  filter(start_station_name == "Buckingham Fountain")

ggplot(data = entire_df)+
  geom_bar(mapping = aes(x=rideable_type, fill=member_casual))+
  labs(
    title = "Starts per Bike-Type: Buckingham Fountain", 
    subtitle = "Dec 2020 - Nov 2021",
    fill = "Rider type",
    x = "Bike-Type"
  )+
  scale_fill_manual(values = c("#339999","#FF0000"))+
  facet_wrap(~member_casual)+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 30),
    panel.grid.major.x = element_blank()
  )

```


The next code chunk was used to generate a bar chart showing the most popular Bike-Type (rideable_type) while faceting by Rider Type (member_casual).


```{r}

library("tidyverse")

entire_df <- read_csv("CompleteCleanedBikeData.csv")

ggplot(data = entire_df)+
  geom_bar(mapping = aes(x=rideable_type, fill=member_casual))+
  labs(
    title = "Starts per Bike-Type", 
    subtitle = "Dec 2020 - Nov 2021",
    fill = "Rider type",
    x = "Bike-Type"
  )+
  scale_fill_manual(values = c("#339999","#FF0000"))+
  facet_wrap(~member_casual)+
  theme_minimal()+
  theme(
    axis.text.x = element_text(angle = 30),
    panel.grid.major.x = element_blank()
  )


```

This concludes the documentation of the R code that was used to find the most popular Bike-Type and the most used stations. All spreadsheets, pivot tables and bar charts were organized into a sildeshow presentation in order to better display the findings.



